---
title: "595 Written Exam"
author: "sbsambado"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

*Show during Bren meeting*
This model is showing population dynamics between hosts and vectors. 
```{r host tick interaction, part 1}
# very basic, simple model to build off of
library(deSolve)
host.tick <- function(t, x, params) {
  S = x[1] # susceptible host
  I = x[2] # infected host
  U = x[3] # susceptible vector
  V = x[4] # infected vector
  
  beta_prime = params[1] # inoculation
  alpha = params[2] # acquisition
  gamma = params[3] # host turnover I > S
  birth_v = params[4] # birth of vectors, or input into the system
  death_v = params[5] # death of vectors
  
  # S + I = N host in system
  dSdt = - (beta_prime*S*V)/(S + I) + gamma*I #change of susceptibles to infected
  dIdt = (beta_prime*S*V)/(S + I) - gamma*I # change of infected to susceptibles
  dUdt = - (alpha*S*I)/(S + I) + birth_v*U - death_v*U + 10
  dVdt = (alpha*S*I)/(S + I) + birth_v*V - death_v*V
 
  # U + V = N vector in system, don't think this would be biologically driving I% but check it out
  #dUdt = - (alpha*U*I)/(U + V) + birth_v*U - death_v*U 
  #dVdt = (alpha*U*I)/(U + V) + birth_v*V - death_v*V
  
  return(list(c(dSdt, dIdt, dUdt, dVdt)))
}

# initial values for state variables
  S0 = 19
  I0 = 1
  U0 = 99
  V0 = 1
initial_values = c(S0, I0, U0, V0)

# initial values for parameters
  beta_prime = 0.7 # inoculation #assuming 70% of infectious contacts turn into infections
  alpha = 0.7  # acquisition #assuming 70% of infectious contacts turn into infections
  gamma = 0.6 # host turnover I > S
  birth_v = 0.5 # birth of vectors
  death_v = 0.6 # death of vectors
parameters = c(beta_prime, alpha, gamma, birth_v, death_v) 

# time series
times = seq(0,100, 0.01)

# run the model
results = lsoda(initial_values, times, host.tick, parameters)
colnames(results) = c('time', 'susceptible host', 'infected host', 'susceptible tick', 'infected tick')
head(results)  

#colors()
Scol <- 'blue'
Icol <- 'red'
Ucol <- 'darkolivegreen'
Vcol <- 'orange'

# plot basic abundances graph
plot(results[, 'time'], results[,'susceptible host'], type = 'l', col = Scol, 
     xlab = 'Time', ylab = 'Abundance', main = 'Population Dyanmics',
     ylim = c(0,101), lty = 2, lwd = 2)
lines(results[,'time'], results[,'infected host'], type = 'l', col = Icol, lty = 1, lwd = 2)
lines(results[,'time'], results[,'susceptible tick'], type = 'l', col = Ucol, lty = 2, lwd = 2)
lines(results[,'time'], results[,'infected tick'], type = 'l', col = Vcol, lty = 1, lwd = 2)
legend('topright', legend = c('Susceptible hosts', 'Infected host', 'Susceptible vectors', 'Infected vectors'),
       col = c(Scol, Icol, Ucol, Vcol), lty = c(2,1,2,1), 
       lwd = 2, cex = 0.9)

# hopefully if I can get psi (input into susceptible vectors) there won't be a crashing Vs population

sd(results[,'infected tick'])
sd(results[,'susceptible tick'])
sd(results[,'infected host'])
sd(results[,'susceptible host'])



H = S0 + I0
T = U0 + V0
# max number of infected hosts at t
#max_infected_host<-(max(results[,'infected host']))/H
# at one point at peak infection, 64% of hosts (N = 13) were infected

# max number of infected hosts at t
#max_infected_tick<-(max(results[,'infected tick']))/T
# at one point at peak infection, 32% of ticks (N = 32) were infected


## what proportion of population had been sick at some point?
plot(x = times, y = ((results[,'infected tick'])/T),
     xlab = 'Time', ylab = 'Proportion of Infected Groups',
     main = 'Proportion of Infected Populations',
     lwd = 2, lty = 1, type = 'l',
     col = Vcol,
     ylim = c(0,1))
lines(x = times, y = ((results[,'infected host'])/H), 
     col = Icol, lwd = 2)
legend('topright', legend = c('Infected Tick', 'Infected Host'),
       col = c(Vcol, Icol), lty = 1, lwd = 2,
       cex = 0.8)

## When was rate of transmission highest?
newinfections_host <- beta_prime*(results[,'susceptible host'])*(results[,'infected host'])
newinfections_tick <- beta_prime*(results[,'susceptible tick'])*(results[,'infected tick'])

plot(x = times, y = newinfections_host,
     type = 'l', col = Icol,
     xlab = 'Time', ylab = 'New Infections', main = 'Rate of Transmission',
     ylim = c(0,2000), xlim = c(0,50))
lines(x = times, y = newinfections_tick, col = Vcol)
legend('topright', legend = c('Infected Tick', 'Infected Host', 'Max Transmission Rate'),
       col = c(Vcol, Icol, 'black'), lty = c(1,1,2), lwd = 2,
       cex = 0.8)
max.newinfections_host <- max(newinfections_host)
max.newinfections_tick <- max(newinfections_tick)

for(i in 1:length(times)) {
  if(newinfections_host[i]==max.newinfections_host){
    H.maxnewinf<- times[i]
  }
}
for(i in 1:length(times)) {
  if(newinfections_tick[i]==max.newinfections_tick){
    T.maxnewinf<- times[i]
  }
}
H.maxnewinf
T.maxnewinf
abline(v = T.maxnewinf, lty = 2, col = Vcol)
abline(v = H.maxnewinf, lty = 2, col = 'black')

```
 Graph 1: Population dynamics of infected and susceptibles for each group
    Makes sense that my highest abundance group is uninfected ticks.
  Graph 2: Proportion of groups are infected
    Makes sense that a higher proportion of hosts are infected than ticks, since there's less hosts     and more opportunities for infectious contacts
  Graph 3: Max transmission occurs in hosts
    Makes sense that max transmission occurs once my hosts reach their peak infection proportions
    show during Bren meeting
    
    
```{r NEEDS WORK change initial values for host.tick model}
## Change initial conditions, like host density to see impact on R0
# make forloop

host.tick.density<- function(t, x, params) {
  S = x[1]
  I = x[2]
  U = x[3]
  V = x[4]
  
  beta_prime = params[1] # inoculation
  alpha = params[2] # acquisition
  gamma = params[3] # host turnover I > S
  birth_v = params[4] # birth of vectors, or input into the system
  death_v = params[5] # death of vectors
  
  # S + I = N host in system
  dSdt = - (beta_prime*S*V) + gamma*I #change of susceptibles to infected
  dIdt = (beta_prime*S*V) - gamma*I # change of infected to susceptibles
  dUdt = - (alpha*S*I) + birth_v*U - death_v*U 
  dVdt = (alpha*S*I) + birth_v*V - death_v*V
 
  return(list(c(dSdt, dIdt, dUdt, dVdt)))
}

Tend = 100
times = seq(0, Tend, by = 1)

# initial values for parameters
  beta_prime = 0.1 # inoculation 
  alpha = 0.1  # acquisition 
  gamma = 1 # host turnover I > S
  birth_v = 0.6 # birth of vectors
  death_v = 0.5 # death of vectors
parameters = c(beta_prime, alpha, gamma, birth_v, death_v) 

# initial values for state variables
  #S0 = 19 changing this
  S0_all <-seq(1,50,1) 
  I0 = 1
  U0 = 200
  V0 = 1


S_longterm = rep(0, length(S0_all))
I_longterm = rep(0, length(S0_all))
U_longterm = rep(0, length(S0_all))
V_longterm = rep(0, length(S0_all))

for(i in length(S0_all)) {
  initial_values = c(S0_all[i], I0, U0, V0)
  
  results = lsoda(initial_values, times, host.tick.density, parameters)
  
  # save final values of state variables
  S_longterm[i] = results[Tend+1,2]
  I_longterm[i] = results[Tend+1,3]
  U_longterm[i] = results[Tend+1,4]
  V_longterm[i] = results[Tend+1,5]
}

plot(S0_all, S_longterm, type = 'l',
     xlab = 'Initial density of Susceptible hosts,S(0)',
     ylab = 'Density of Susceptibles Remaining',
     main = 'Change initial density of S0',
     ylim = c(0, 100))
lines(S0_all, S0_all, type = 'l', col = Scol,lty = 2) # 1:1 line; final density of S = S0
abline(v = gamma/beta_prime, col = 'red') # threshold density of S for pathogen invasion
legend("topright", legend = c('1:1 line','S0', 'Threshold Density'),
       col = c('blue', 'black', 'red'), lty = c(2,1,1))

```


```{r change initial values of beta prime NEED TO FIX}
## Change initial conditions, like beta_prime to see impact on R0
# make forloop

host.tick.density<- function(t, x, params) {
  S = x[1]
  I = x[2]
  U = x[3]
  V = x[4]
  
  beta_prime = params[1] # inoculation
  alpha = params[2] # acquisition
  gamma = params[3] # host turnover I > S
  birth_v = params[4] # birth of vectors, or input into the system
  death_v = params[5] # death of vectors
  
  # S + I = N host in system
  dSdt = - (beta_prime*S*V) + gamma*I #change of susceptibles to infected
  dIdt = (beta_prime*S*V) - gamma*I # change of infected to susceptibles
  dUdt = - (alpha*S*I) + birth_v*U - death_v*U 
  dVdt = (alpha*S*I) + birth_v*V - death_v*V
 
  return(list(c(dSdt, dIdt, dUdt, dVdt, beta_prime)))
}

Tend = 100
times = seq(0, Tend, by = 1)

# initial values for parameters
  beta_prime_all <- seq(.001, .999, .01)
  #beta_prime = 0.1 # inoculation 
  alpha = 0.1  # acquisition 
  gamma = 1 # host turnover I > S
  birth_v = 0.6 # birth of vectors
  death_v = 0.5 # death of vectors
parameters = c(beta_prime, alpha, gamma, birth_v, death_v) 

# initial values for state variables
  S0 = 19 
  #S0_all <-seq(1,50,1) 
  I0 = 1
  U0 = 200
  V0 = 1

z = 1:100
S_longterm = rep(0, length(z))
I_longterm = rep(0, length(z))
U_longterm = rep(0, length(z))
V_longterm = rep(0, length(z))

for(i in length(z)) {

  parameters = c(beta_prime_all[i], alpha, gamma, birth_v, death_v) 

  results = lsoda(initial_values, times, host.tick.density, parameters)
  
  # save final values of state variables
  S_longterm[i] = results[Tend+1,2]
  I_longterm[i] = results[Tend+1,3]
  U_longterm[i] = results[Tend+1,4]
  V_longterm[i] = results[Tend+1,5]
  beta_prime_longterm[i] = results[Tend+1,6]
}

plot(beta_prime_all,beta_prime_longterm, type = 'l' )
plot(S0_all, S_longterm, type = 'l',
     xlab = 'Initial density of Susceptible hosts,S(0)',
     ylab = 'Density of Susceptibles Remaining',
     main = 'Change initial density of S0',
     ylim = c(0, 250))
lines(S0_all, S0_all, type = 'l', col = Scol,lty = 2) # 1:1 line; final density of S = S0
abline(v = gamma/beta_prime, col = 'red') # threshold density of S for pathogen invasion
legend("topright", legend = c('1:1 line','S0', 'Threshold Density'),
       col = c('blue', 'black', 'red'), lty = c(2,1,1))

```

```{r for loop of host tick interaction model pt. 1 NEED TO FIX}

## need to fix my ODEs here

# initial values for parameters  
  beta_prime = 0.9 # inoculation
  alpha = 0.8  # acquisition
  gamma = 0.7 # host turnover I > S
  birth_v = 0.3 # birth of vectors
  death_v = 0.5 # death of vectors

# inital values for state variables
  H = 30
  T = 100
  
  S0 = 29
  I0 = 1
  U0 = 99
  V0 = 1

# set up storage variables
tset <- seq(0, 60, 2000)
H.simu1 <- NaN*tset; H.simu1[1] <- H
T.simu1 <- NaN*tset; T.simu1[1] <- T
S.simu1 <- NaN*tset; S.simu1[1] <- S0
I.simu1 <- NaN*tset; I.simu1[1] <- I0
U.simu1 <- NaN*tset; U.simu1[1] <- U0
V.simu1 <- NaN*tset; V.simu1[1] <- V0

for(i in 2:length(tset)) {
  dt <- tset[i] - tset[i -1]
  # rename to shorten commands to make ODEs easier
  S <- S.simu1[i-1]
  I <- I.simu1[i-1]
  U <- U.simu1[i-1]
  V <- V.simu1[i-1]
  
  dS <- (- (beta_prime*S*V)/(S + I) + gamma*I)*dt
  dI <- ( (beta_prime*S*V)/(S + I) - gamma*I)*dt
  dU <- (- (alpha*S*I)/(S + I) + birth_v*U - death_v*U)*dt
  dV <- ((alpha*S*I)/(S + I) - death_v*U)*dt
  
  dH <- dS + dI
  dT <- dU + dV
  
  H.simu1[i] <- H +dH
  T.simu1[i] <- T +dT
  S.simu1[i] <- S +dS
  I.simu1[i] <- I +dI
  U.simu1[i] <- U +dU
  V.simu1[i] <- V +dV
}


```



*Show during Bren meeting?*
This model is showing the influence of transmission mode inputs as ticks transition from questing to feeding.
```{r host tick interactions part 2, how to transfer from questing to feeding}
q.to.f <- function(t, x, params) {
  Xqii = x[1] # infected stays infected
  Xqsi = x[2] # susceptible becomes infected
  Xqss = x[3] # susceptible stays susceptible
  
  
  lambda= params[1] # inoculation
  alpha = params[2] # acquisition via infected blood meal
  sigma = params[3] # acquisition via cofeeding
  psi = params[4] # input of new vectors into questing X
  omega = params[5] # death of vectors
  
  # S + I = N host in system
  dXqiidt = psi - omega*Xqii - lambda*Xqii # infected stays infected
  dXqsidt = psi - omega*Xqsi - lambda*Xqsi*alpha - lambda*Xqsi*sigma # susceptible becomes infected
  # if sigma/alpha 0, selects 1 route of transmission
  dXqssdt = psi - omega*Xqss - lambda*Xqss # susceptible stays susceptible
  
  return(list(c(dXqiidt, dXqsidt, dXqssdt)))
}

Xqii0 = 10
Xqsi0 = 30
Xqss0 = 60

## set values
# state variables
initial_values = c(Xqii0, Xqsi0, Xqss0)

## all transmission routes are equal
# parameters
  lambda= .7 # contact rate
  alpha_equal = .01 # acquisition via infected blood meal
  sigma_equal = .01 # acquisition via cofeeding
  psi = 50 # input of new vectors into questing X
  omega = .5 # death of vectors

parameters_equal = c(lambda, alpha_equal, sigma_equal, psi, omega) 

times = seq(0,5, 0.01)

results_equal = lsoda(initial_values, times, q.to.f, parameters_equal)
colnames(results_equal) = c('time', 'I to I', 'S to I', 'S to S')
head(results_equal)  

## HT greater >>
# parameters
  lambda= .7 # contact rate
  alpha_greater = .7 # acquisition via infected blood meal
  sigma_equal = .01 # acquisition via cofeeding
  psi = 50 # input of new vectors into questing X
  omega = .5 # death of vectors

parameters_HT = c(lambda, alpha_greater, sigma_equal, psi, omega) 

times = seq(0,5, 0.01)

results_HT = lsoda(initial_values, times, q.to.f, parameters_HT)
colnames(results_HT) = c('time', 'I to I', 'S to I', 'S to S')
head(results_HT)  

## co-feeding greater
# parameters
  lambda= .7 # contact rate
  alpha_small = .1 # acquisition via infected blood meal
  sigma_greater = .4 # acquisition via cofeeding
  psi = 50 # input of new vectors into questing X
  omega = .5 # death of vectors

parameters_CF = c(lambda, alpha_small, sigma_greater, psi, omega) 

times = seq(0,5, 0.01)

results_CF = lsoda(initial_values, times, q.to.f, parameters_CF)
colnames(results_CF) = c('time', 'I to I', 'S to I', 'S to S')
head(results_CF)  


qiicol <-"blue"
qsicol<- 'red'
qsscol <- 'purple'

# use main = c('', paste())
# alternative title Transitions from Questing to Feeding. 
par(mfrow = c(3,1), mar = c(2,4,3,1))
# equal

## try to find max of susceptible ticks going into next system


plot(results_equal[, 'time'], results_equal[,'I to I'], type = 'l', col = qiicol, 
     xlab = 'Time', ylab = 'Abundance',main = c('Transmission inputs: Non-systemic = Systemic'),cex.main = 1,
     ylim = c(0,100), lty = 1, lwd = 2)
lines(results_equal[,'time'], results_equal[,'S to I'], type = 'l', col = qsicol, lty = 1, lwd = 2)
lines(results_equal[,'time'], results_equal[,'S to S'], type = 'l', col = qsscol, lty = 1, lwd = 2)
legend('bottomright', legend = c('I to I', 'S to I', 'S to S'),
       col = c('blue', 'purple', 'red'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 1, horiz = TRUE)

# greater HT
plot(results_HT[, 'time'], results_HT[,'I to I'], type = 'l', col = qiicol, 
     xlab = 'Time', ylab = 'Abundance', main = 'Transmission inputs: Non-systemic << Systemic',cex.main = 1,
     ylim = c(0,100), lty = 1, lwd = 2)
lines(results_HT[,'time'], results_HT[,'S to I'], type = 'l', col = qsicol, lty = 1, lwd = 2)
lines(results_HT[,'time'], results_HT[,'S to S'], type = 'l', col = qsscol, lty = 1, lwd = 2)
legend('bottomright', legend = c('I to I', 'S to I', 'S to S'),
       col = c('blue', 'purple', 'red'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 1, horiz = TRUE)

# greater CF
plot(results_CF[, 'time'], results_CF[,'I to I'], type = 'l', col = qiicol, 
     xlab = 'Time', ylab = 'Abundance', main =  'Transmission inputs: Non-systemic > Systemic',cex.main = 1, 
     ylim = c(0,100), lty = 1, lwd = 2)
lines(results_CF[,'time'], results_CF[,'S to I'], type = 'l', col = qsicol, lty = 1, lwd = 2)
lines(results_CF[,'time'], results_CF[,'S to S'], type = 'l', col = qsscol, lty = 1, lwd = 2)
legend('bottomright', legend = c('I to I', 'S to I', 'S to S'),
       col = c('blue', 'purple', 'red'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 1, horiz = TRUE)


```
Systemic feeding is tick acquiring infection horizontally from a blood meal. (HT)
Non-Systemic feeding is tick acquiring infection from a co-feeding event. (CF)
  Graph 1: HT = CF
  Graph 2: HT >> CF (most likely event)
  Graph 3: HF < CF
I need to analyze the results better (particularly when and how many S are there in the system.

What I really want to know is how many ticks leave this phase susceptible, which would decrease transmission at the next phase. Or how many ticks leave this phase infected, which would increase transmission at the next phase, maybe increase the role of co-feeding.


*Show during Bren meeting?*
This model is trying to understand how temperature influences the transition of tick phases.
The idea is, the warmer the temperature, the quicker ticks transition, which will ultimately alter tick populations and what phases are active (i.e. disease risk)
```{r tick phase model, playing around with temp sensitive parameters}
all.phases <- function(t, x, params) {
  # 3 classes from questing to feeding
  Xq = x[1] # infected stays infected
  Xf = x[2]
  Xe = x[3]
  
  lambda = params[1]
  phi = params[2]
  tau = params[3]
  
  psi = params[4]
  omega = params[5]
  
  # S + I = N host in system
  
  # questing to feeding
  dXqdt = psi - omega*Xq - lambda*Xq # infected stays infected
  
  
  # feeding to engorged
  dXfdt = lambda*Xq - omega*Xf - phi*Xf
  
  
  # engorged to molted to next X
  dXedt = phi*Xf - omega*Xe - tau*Xe

  
  return(list(c(dXqdt,dXfdt,dXedt)))
}

Xq0 = 100
Xf0 = 60
Xe0 = 50

## set values
# state variables
initial_values = c(Xq0,Xf0,Xe0)

# parameters_equal
lambda = .5 # contact rate
  phi = .5 # drop off rate
  tau = .5 # molting rate
  
  psi = 10 # input of new vectors into questing X
  omega = .5 # death of vectors
  
parameters_equal = c(lambda, phi, tau, psi, omega) 

times = seq(0,5, 0.1)

results_equal = lsoda(initial_values, times, all.phases, parameters_equal)
colnames(results_equal) = c('time', 'Xq','Xf','Xe')
head(results_equal)  

# greater parameters
lambda_greater = .8 # contact rate
  phi_greater = .8 # drop off rate
  tau_greater = .8 # molting rate
  
  psi = 10 # input of new vectors into questing X
  omega = .5 # death of vectors
  
parameters_greater = c(lambda_greater, phi_greater, tau_greater, psi, omega) 

times = seq(0,5, 0.1)

results_greater = lsoda(initial_values, times, all.phases, parameters_greater)
colnames(results_greater) = c('time', 'Xq','Xf','Xe')
head(results_greater)  

# lower parameters
lambda_lower = .3 # contact rate
  phi_lower = .3 # drop off rate
  tau_lower = .3 # molting rate
  
  psi = 10 # input of new vectors into questing X
  omega = .5 # death of vectors
  
parameters_lower = c(lambda_lower, phi_lower, tau_lower, psi, omega) 

times = seq(0,5, 0.1)

results_lower = lsoda(initial_values, times, all.phases, parameters_lower)
colnames(results_lower) = c('time', 'Xq','Xf','Xe')
head(results_lower)  

#par(mfrow = c(3,1))
plot(results_greater[, 'time'], results_greater[,'Xq'], type = 'l', col = 'red', 
     xlab = 'Time', ylab = 'Abundance',main = 'Tick population dynamics, temp parameters increased',
     ylim = c(0,100), lty = 1, lwd = 2)
lines(results_greater[,'time'], results_greater[,'Xf'], type = 'l', col = 'green', lty = 1, lwd = 2)
lines(results_greater[,'time'], results_greater[,'Xe'], type = 'l', col = 'blue', lty = 1, lwd = 2)
legend('topright', legend = c('questing','feeding','engorged'),
       col = c('red', 'blue', 'green'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 0.7, horiz = TRUE)
plot(results_equal[, 'time'], results_equal[,'Xq'], type = 'l', col = 'red', 
     xlab = 'Time', ylab = 'Abundance',main = 'Tick population dynamics, temp. parameters base level',
     ylim = c(0,100), lty = 1, lwd = 2)
lines(results_equal[,'time'], results_equal[,'Xf'], type = 'l', col = 'green', lty = 1, lwd = 2)
lines(results_equal[,'time'], results_equal[,'Xe'], type = 'l', col = 'blue', lty = 1, lwd = 2)
legend('topright', legend = c('questing','feeding','engorged'),
       col = c('red', 'blue', 'green'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 0.7, horiz = TRUE) 
plot(results_lower[, 'time'], results_lower[,'Xq'], type = 'l', col = 'red', 
     xlab = 'Time', ylab = 'Abundance',main = 'Tick population dynamics, temp. parameters decreased', ylim = c(0,100), lty = 1, lwd = 2)
lines(results_lower[,'time'], results_lower[,'Xf'], type = 'l', col = 'green', lty = 1, lwd = 2)
lines(results_lower[,'time'], results_lower[,'Xe'], type = 'l', col = 'blue', lty = 1, lwd = 2)
legend('topright', legend = c('questing','feeding','engorged'),
       col = c('red', 'blue', 'green'), lty = c(1,1,1), 
       lwd = c(2,2,2), cex = 0.7, horiz = TRUE)
```
I am focusing on the abundance of questing activity since that is associated with disease risk and transmission.
  Graph 1: increased temperature, increase parameter rate
  Graph 2: current temperature, current parameter rate
  Graph 3: decreased temperature, decreased parameter rate (unlikely scenario, maybe should do a scenario where it gets too hot)

*5/13 model*
This model has reduced amount of state variables
```{r}
reduced <- function(t, x, params) {
  Lqi = x[1]
  Lqs = x[2]
  Lfi = x[3]
  Lfs = x[4]
  
  Nqi = x[5]
  Nqs = x[6]
  Nfi = x[7]
  Nfs = x[8]
  
  Aqi = x[9]
  Aqs = x[10]
  Afi = x[11]
  Afs = x[12]
  
  t = params[1]
  k = params[2]
  b = params[3] # contact rate
  p = params[4]
  y = params[5] # birth rate
  H = params[6]
  Hi = params[7]
  Hs = params[8]
  v = params[9] # no VT
  
  dLqidt = Afi*y*p - b*H*Lqi
  dLqsdt = Afi*y*w + y*w*Afs - b*Hi*Lqi - b*Hs*Lqs
  dLfidt = b*Hi*Lqi + b*Hs*Lqs - t*k*Lfi
  dLfsdt = b*Hs*Lqs - t*k*Lfs
  
  dNqidt = t*k*Lfi - b*H*Nqi
  dNqsdt = t*k*Lfs - b*Hs*Nqs - b*Hi*Nqs
  dNfidt = b*H*Nqi + b*Hi*Nqs - t*k*Nfi
  dNfsdt = b*Hs*Nqs - t*k*Nfs 
  
  dAqidt = t*k*Nfi - b*H*Aqi
  dAqsdt = t*k*Nfs - b*Hs*Aqi - b*Hi*Aqs
  dAfidt =  b*H*Aqi + b*Hi*Aqs - y*p*Afi - y*w*Afi
  dAfsdt = b*Hs*Aqs - y*w*Afs
    
  return(list(c(dLqidt, dLqsdt, dLfidt, dLfidt,
                dNqidt, dNqsdt, dNfidt, dNfidt,
                dAqidt, dAqsdt, dAfidt, dAfidt
                )))
}

  Lqi0 = 30
  Lqs0 = 70
  Lfi0 = 30
  Lfs0 = 50
  
  Nqi0 = 20
  Nqs0 = 30
  Nfi0 = 15
  Nfs0 = 25
  
  Aqi0 = 30
  Aqs0 = 10
  Afi0 = 25
  Afs0 = 5
  
initial_values = c(Lqi0, Lqs0, Lfi0, Lfs0,
                   Nqi0, Nqs0, Nfi0, Nfs0,
                   Aqi0, Aqs0, Afi0, Afs0)

t = 1.5 # temp sensitive
k = .4 # survival
b = .8 # contact rate
p = .7 # VT
y = .9 # birth rate
H = 40 # hi + hs
Hi = 25
Hs = 15
v = .3 # no VT

parameters = c(t,k, b, p, y, H, Hi, Hs,v)

times = seq(0,5, by = 0.01)

results = lsoda(initial_values, times, reduced, parameters)
colnames(results) = c('time', 'Lqi','Lqs','Lfi','Lfs',
                       'Nqi','Nqs','Nfi','Nfs',
                       'Aqi','Aqs','Afi','Afs')
#head(results)
plot(results[,'time'], results[,'Lqi'], type = 'l', col = 'red')
lines(results[,'time'], results[,'Nqi'], type = 'l', col = 'blue')
lines(results[,'time'], results[,'Aqi'], type = 'l', col = 'green')
# uninfected
lines(results[,'time'], results[,'Lqs'], type = 'l', col = 'red', lty = 2)
lines(results[,'time'], results[,'Nqs'], type = 'l', col = 'blue', lty = 2)
lines(results[,'time'], results[,'Aqs'], type = 'l', col = 'green', lty = 2)



```


{r all phase transitions within one tick lifes stage LOL}
all.phases <- function(t, x, params) {
  # 3 classes from questing to feeding
  Xqi = x[1] # infected stays infected
  Xqs.a = x[2] # susceptible becomes infected
  Xqs.b = x[3] # susceptible stays susceptible
  
  # 2 classes from feeding to engorged
  Xfi = x[4]
  Xfs = x[5]
  # 2 classes from engorged to molted
  Xei = x[6]
  Xes = x[7]
    
  lambda = params[1] # inoculation
  alpha = params[2] # acquisition via infected blood meal
  sigma = params[3] # acquisition via cofeeding
  psi = params[4] # input of new vectors into questing X
  omega = params[5] # death of vectors
  phi = params[6] # dropping off rate
  tau = params[7] # molting rate
  
  # S + I = N host in system
  
  # questing to feeding
  dXqidt = psi - omega*Xqi - lambda*Xqi # infected stays infected
  
  # this line needs help, 
  #which do I assign death to? Do I combine both routes for si here?
  dXqs.adt = psi - omega*Xqs.a - lambda*Xqs.a*alpha - omega*Xqs.a -lambda*Xqs.a*sigma # susceptible becomes infected
  # if sigma/alpha 0, selects 1 route of transmission
  
  dXqs.bdt = psi - omega*Xqs.b - lambda*Xqs.b # susceptible stays susceptible
  
  # feeding to engorged
  dXfidt = lambda*Xqi + lambda*Xqs.a*alpha + lambda*Xqs.a*sigma - omega*Xfi - phi*Xfi
  dXfsdt = lambda*Xqs.b - omega*Xfs - phi*Xfs
  
  # engorged to molted to next X
  dXeidt = phi*Xfi - omega*Xei - tau*Xei
  dXesdt = phi*Xfs - omega*Xes - tau*Xes
  
  return(list(c(dXqidt, dXqs.adt, dXqs.bdt, dXfidt,dXfsdt,dXeidt,dXesdt)))
}

Xqi0 = 1
Xqs.a0 = 3
Xqs.b0 = 6

Xfidt0 = 4
Xfsdt0 = 6

Xeidt0 = 4
Xesdt0 = 6

## set values
# state variables
initial_values = c(Xqi0, Xqs.a0, Xqs.b0,
                   Xfidt0,Xfsdt0,
                   Xeidt0,Xesdt0)

# parameters
  lambda= .5 # contact rate
  alpha = .3 # acquisition via infected blood meal
  sigma = .1 # acquisition via cofeeding
  psi = 50 # input of new vectors into questing X
  omega = .5 # death of vectors
  phi = .5
  tau = .6
parameters = c(lambda, alpha, sigma, psi, omega, phi, tau) 

times = seq(0,30, 0.01)

results = lsoda(initial_values, times, all.phases, parameters)
colnames(results) = c('time', 'Xqi', 'Xqs.a', 'Xqs.b','Xfi','Xfs','Xei','Xes')
head(results)  


plot(results[, 'time'], results[,'Xqi'], type = 'l', col = 'bisque1', 
     xlab = 'Time', ylab = 'Abundance',main = 'Tick population dynamics',
     ylim = c(0,60), lty = 1, lwd = 2)
lines(results[,'time'], results[,'Xqs.a'], type = 'l', col = 'bisque2', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Xqs.b'], type = 'l', col = 'bisque3', lty = 2, lwd = 1)
lines(results[,'time'], results[,'Xfi'], type = 'l', col = 'aquamarine1', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Xfs'], type = 'l', col = 'aquamarine2', lty = 2, lwd = 1)
lines(results[,'time'], results[,'Xei'], type = 'l', col = 'lightgray', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Xes'], type = 'l', col = 'gray', lty = 2, lwd = 1)
legend('topright', legend = c('Xqi', 'Xqs.a', 'Xqs.b','Xfi','Xfs','Xei','Xes'),
       col = c('bisque1', 'bisque2', 'bisque3', 'aquamarine1', 'aquamarine2', 'lightgray', 'gray'), lty = c(1,1,2,1,2,1), 
       lwd = c(2,2,1,2,1,2,1), cex = 0.7, horiz = TRUE)
```


```{r add eggs LOL capital of crazytown}
eggstoadults <- function(t, x, params) {
  
  # LARVAE
  # omit 3 classes from questing to feeding
  Lqv = x[1] # infected stays infected
  Lqu = x[2]
  #Lqs.a = x[2] # susceptible becomes infected
  #Lqs.b = x[3] # susceptible stays susceptible
  
  # 2 classes from feeding to engorged
  Lfi = x[3]
  Lfs = x[4]
  # 2 classes from engorged to molted
  Lei = x[5]
  Les = x[6]
  
  #NYMPHS
  Nqi = x[7] # infected stays infected
  Nqs = x[8] # susceptible becomes infected
  #Nqs.b = x[10] # susceptible stays susceptible
  
  # 2 classes from feeding to engorged
  Nfi = x[9]
  Nfs = x[10]
  # 2 classes from engorged to molted
  Nei = x[11]
  Nes = x[12]
  
  
  #ADULTS
  Aqi = x[13] # infected stays infected
  Aqs = x[14]
  
  # 2 classes from feeding to engorged
  Afi = x[15]
  Afs = x[16]
  
  # eggs
  Gv = x[17]
  Gu = x[18]
  
  # hosts
  Hi = x[19]
  Hs = x[20]
  
  Di = x[21]
  Ds = x[22]
    
  kappa = params[1] # survival prop to next life stage
  rho = params[2] # filial infection prevalence (fip)
  rho_prime = params[3] # no fip
  beta = params[4] # transmission between host and questing ticks (contact x prob infect)
  phi = params[5] # dropping off rate
  tau = params[6] # molting rate
  delta = params[7] # birth rate
  tot = params[8] # total ovarial transmission (tot)
  tot_prime = params[9] # no tot
  B = params[10] #contact rate
  q = params[11] # prob tick infected 
  gamma = params[12] # recovery rate of host
  

  
  # questing to feeding
  dLqvdt = kappa*rho*Gv - beta*(Hs + Hi)*Lqv 
  dLqudt = kappa*rho_prime*Gv + kappa*Gu - beta*Hi*Lqu - beta*Hs*Lqu
  
  # feeding to engorged
  dLfidt = beta*(Hs + Hi)*Lqv + beta*Hi*Lqu - phi*Lfi
  dLfsdt = beta*Hs*Lqu - phi*Lfs
  
  # engorged to molted to next X
  dLeidt = phi*Lfi - kappa*tau*Lei 
  dLesdt = phi*Lfs - kappa*tau*Les
  
  ## NYMPHS
  # questing to feeding
  dNqidt = kappa*tau*Lei - beta*(Hs + Hi)*Nqi 
  dNqsdt = kappa*tau*Les - beta*Hi*Nqs - beta*Hs*Nqs 
  
  # feeding to engorged
  dNfidt = beta*H*Nqi + beta*Hi*Nqs - phi*Nfi
  dNfsdt = beta*Hs*Nqs - phi*Nfs
  
  # engorged to molted to next X
  dNeidt = phi*Nfi - kappa*tau*Nei 
  dNesdt = phi*Nfs - kappa*tau*Nes
  
  ## Adults
  # questing to feeding
  dAqidt = kappa*tau*Nei - beta*(Ds+Di)*Aqi 
  dAqsdt = kappa*tau*Nes - beta*Di*Aqs - beta*Ds*Aqs 
  
  # feeding to birthing
  dAfidt = beta*(Ds + Di)*Aqi + beta*Di*Aqs - delta*tot*Afi - delta*tot_prime*Afi
  dAfsdt = beta*Ds*Aqs - delta*Afs
  
 # birthing/eggs
  dGvdt = delta*tot*Afi - kappa*rho*Gv - kappa*rho_prime*Gv
  dGudt = delta*tot*Afs + delta*tot_prime*Afi - kappa*Gu
  
# host populations
  
  dHsdt = - ((Hs*(B*q)*Lqi*Nqi)/(Hs + Hi)) + gamma*Hi # assuming contact rate and inf prob are = for L and N but need to change
  dHidt = ((Hs*(B*q)*Lqi*Nqi)/(Hs + Hi)) - gamma*Hi
  
  dDsdt = - ((Ds*(B*q)*Aqi)/(Di + Ds)) + gamma*Di
  dDidt =  ((Ds*(B*q)*Aqi)/(Di + Ds)) - gamma*Di
  
   # S + I = N host in system
  dHdt = dHidt + dHsdt # rodent hosts
  dDdt = dDidt + dDsdt # deer hosts
  
  return(list(c(dLqvdt, dLqudt, dLfidt, dLfsdt, dLeidt, dLesdt,
                dNqidt, dNqsdt, dNfidt, dNfsdt, dNeidt, dNesdt,
                dAqidt, dAqsdt, dAfidt, dAfsdt, 
                dGvdt, dGudt,
                dHsdt, dHidt,
                dDsdt, dDidt)))
}

Lqv0 = 10
Lqu0 = 10
Lfi0 = 10
Lfs0 = 10
Lei0 = 10
Les0 = 10
                
Nqi0 = 5 
Nqs0 = 5
Nfi0 = 5
Nfs0 = 5
Nei0 = 5
Nes0 = 5

Aqi0 = 1
Aqs0 = 1
Afi0 = 1
Afs0 = 1

Gv0 = 20
Gu0 = 20

Hs0 = 5
Hi0 = 5
Ds0 = 2
Di0 = 2

## set values
# state variables
initial_values = c(Lqv0,Lqu0,Lfi0,Lfs0,Lei0,Les0,
                Nqi0, Nqs0,Nfi0,Nfs0,Nei0,Nes0,
                Aqi0,Aqs0,Afi0,Afs0,
                Gv0,Gu0,
                Hs0,Hi0,
                Ds0,Di0)

# parameters
  kappa = .9 # survival prop to next life stage
  rho = .6 # filial infection prevalence (fip)
  rho_prime = .4 # no fip
  beta = .9 # transmission between host and questing ticks (contact x prob infect)
  phi = .9 # dropping off rate
  tau = .6 # molting rate
  delta =  .3 # birth rate
  tot = .3 # total ovarial transmission (tot)
  tot_prime = .7 # no tot
  B = .8#contact rate
  q = .5 # prob tick infected 
  gamma = .4 # recovery rate of host
  
parameters = c(kappa, rho, rho_prime, beta, phi,
               tau, delta, tot, tot_prime, B, q, gamma) 

times = seq(0,50, 0.1)

results = lsoda(initial_values, times, eggstoadults, parameters)
colnames(results) = c('time', 'Lqv', 'Lqu', 'Lfi', 'Lfs', 'Lei', 'Les',
                      'Nqi', 'Nqs','Nfi', 'Nfs', 'Nei', 'Nes', 
                      'Aqi', 'Aqs', 'Afi','Afs',
                      'Gv', 'Gu',
                      'Hs', 'Hi',
                      'Ds', 'Di')
head(results)  
#View(results)

# plot infected
plot(results[, 'time'], results[,'Lqv'], type = 'l', col = 'red', 
     xlab = 'Time', ylab = 'Abundance', main = 'Infected Tick Life Stages',
     ylim = c(0,50), xlim = c(0,15),
     lty = 1, lwd = 2)
lines(results[,'time'], results[,'Nqi'], type = 'l', col = 'blue', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Aqi'], type = 'l', col = 'green', lty = 1, lwd = 2)



plot(results[, 'time'], results[,'Lqv'], type = 'l', col = 'black',
     xlab = 'time', ylab = 'abundance',
     main = 'crazyland',ylim = c(0,100), lty = 1, lwd = 1)
lines(results[,'time'], results[,'Lqu'], type = 'l', col = 'grey', lty = 1, lwd = 1)
lines(results[,'time'], results[,'Lfi'], type = 'l', col = 'lightblue1', lty = 2, lwd = 1)
lines(results[,'time'], results[,'Lfs'], type = 'l', col = 'lightblue2', lty = 1, lwd = 1)
lines(results[,'time'], results[,'Lei'], type = 'l', col = 'lightblue3', lty = 2, lwd = 1)
lines(results[,'time'], results[,'Les'], type = 'l', col = 'lightblue4', lty = 1, lwd = 1)
lines(results[,'time'], results[,'Nqi'], type = 'l', col = 'lightsalmon', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Nqs'], type = 'l', col = 'lightsalmon1', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Nfi'], type = 'l', col = 'lightsalmon2', lty = 2, lwd = 2)
lines(results[,'time'], results[,'Nfs'], type = 'l', col = 'lightsalmon3', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Nei'], type = 'l', col = 'lightsalmon4', lty = 2, lwd = 2)
lines(results[,'time'], results[,'Nes'], type = 'l', col = 'lightpink', lty = 1, lwd = 2)
lines(results[,'time'], results[,'Aqi'], type = 'l', col = 'plum', lty = 1, lwd = 3)
lines(results[,'time'], results[,'Aqs'], type = 'l', col = 'plum1', lty = 1, lwd = 3)
lines(results[,'time'], results[,'Afi'], type = 'l', col = 'plum2', lty = 2, lwd = 3)
lines(results[,'time'], results[,'Afs'], type = 'l', col = 'plum3', lty = 1, lwd = 3)


```



```{r absurd number of state variables; NEED TO FIX}
# types of competent and non-competent hosts, need to figure out this bit
S = 10
I = 1
H = S + I
# H1 = 
# H2 =
# H3 =
# H_types = H1 + H2 + H3

# susceptible hosts
H_s1 = 5 # highly competent # NEFU
H_s2 = 7 # lowly comptenty  # PERO
H_s3 = 9 # not competent    # SCOC
S = H_s1 + H_s2 +H_s3

# infected hosts  
H_i1 = 4 # highly competent # NEFU
H_i2 = 0 # lowly comptenty  # PERO
H_i3 = 0 #not competent    # SCOC # this is a weird one
I = H_i1 + H_i2 +H_i3 


# susceptible ticks
Ls = 12
Ns = 8
As = 2
Vs = Ls + Ns + As

# infected ticks
Li = 1
Ni = 0
Ai = 0
Vi = Li + Ni + Ai

# tick life stages
X = Vs + Vi
L = 10
N = 8
A = 2
X_ls = L + N + A

# infected questing ticks 
Lqi = 1
Nqi = 11
Aqi = 9
q = Lqi + Nqi + Aqi 

```

```{r vary parameters, NEED TO FIX}
# tick natural mortality

wl = .1
wn = .2
wa = .2
w = wl + wn + wa

# host recovery rate 
gamma = .9

# tick input to each life stage
psi = X_ls*w # need to figure this out

 
# contact rate between tick and host
beta_l = .05
beta_n = .03
beta_a = .01
beta_x = beta_l + beta_n + beta_a

# tick acquisition rate of pathogen ( host to tick )
alpha = (I * beta_x * q)/H

# I want to divie this up into contact between different types of competent hosts eventually
# beta_l1 =
#   beta_l2 =
#   beta_l3 =
# beta_n1 =
#   beta_n2 =
#   beta_n3 =
# beta_a1 =
#   beta_a2 =
#   beta_a3 =
  # infectious contact rate between tick and host
beta = beta_x * q

# host inoculation rate of patogen ( tick to host )
beta_prime = (H * beta * X)/ H


```

This is trash, fix or delete
```{r toy model LOL trash}
toy2.0 <- function(t, x, params) {
  S = x[1] # susceptible host
  I = x[2] # infected host
  U = x[3] # susceptible vector
  V = x[4] # infected vector
  
  beta_prime = params[1] # inoculation
  alpha = params[2] # acquisition
  gamma = params[3] # host turnover I > S
  birth_v = params[4] # birth of vectors, or input into the system
  death_v = params[5] # death of vectors
  
  # S + I = N host in system
  dSdt = - beta_prime + gamma*I #change of susceptibles to infected
  dIdt = beta_prime - gamma*I # change of infected to susceptibles
  dUdt = - alpha - death_v*Vs + psi # need to switch birth_v to psi
  dVdt = alpha + birth_v*Vs - death_v*Vi # need to switch birth_v to psi
 
  # U + V = N vector in system, don't think this would be biologically driving I% but check it out
  #dUdt = - (alpha*U*I)/(U + V) + birth_v*U - death_v*U 
  #dVdt = (alpha*U*I)/(U + V) + birth_v*V - death_v*V
  
  return(list(c(dSdt, dIdt, dUdt, dVdt)))
  

}
S0 = 19
I0 = 1
U0 = 99
V0 = 1

initial_values = c(S0, I0, U0, V0)


  # beta_prime = 0.9 # inoculation
  # alpha = 0.8  # acquisition
  # gamma = 0.7 # host turnover I > S
  # birth_v = 0.2 # birth of vectors
  # death_v = 0.5 # death of vectors

parameters = c(beta_prime, alpha, gamma, birth_v, death_v) 

times = seq(0,30, 0.01)

results = lsoda(initial_values, times, toy2.0, parameters)
colnames(results) = c('time', 'susceptible host', 'infected host', 'susceptible tick', 'infected tick')
head(results)  

plot(results[, 'time'], results[,'susceptible host'], type = 'l', col = 'blue', 
     xlab = 'time', ylab = 'abundance', ylim = c(0,101), lty = 1, lwd = 1)
lines(results[,'time'], results[,'infected host'], type = 'l', col = 'blue', lty = 2, lwd = 2)
lines(results[,'time'], results[,'susceptible tick'], type = 'l', col = 'red', lty = 1, lwd = 1)
lines(results[,'time'], results[,'infected tick'], type = 'l', col = 'red', lty = 2, lwd = 2)
legend('bottomright', legend = c('Susceptible hosts', 'Infected host', 'Susceptible vectors', 'Infected vectors'),
       col = c('blue', 'blue', 'red', 'red'),lty = c(1,2,1,2), cex = 0.9,
       lwd = c(1,2,1,2))
```


