---
title: "Giardina model"
author: "sbsambado"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Variation in composition of communities of ticks and hosts will influence tick infection prevalence and disease risk 

number of blood meals
  ml, i = number of blood meals for larvae on i host
  mn, i = number of blood meals for nymphs on i host
  
i = species of host
  1 = mice
  2 = wood rat
  3 = lizard

N = number of i species

B = number of ticks on host 
  Bl,i = number of larvae on i species
  Bn,i = number of nymphs on i species


```{r number of blood meals by species and tick life stage}
# inital values
l10 = 2
n10 = 1

l20 = 2
n20 = 1
  
l30 = 2
n30 = 1

# parameters

# density of species
N1 = 50 #mice
N2 = 25 # NEFU
N3 = 30 # lizards

# contact rate of species and life stage of ticks

# contact rate between i species and larval ticks
Bl1 = 9
Bl2 = 6
Bl3 = 15
 
# contact rate between i species and nymphal ticks
Bn1 = 4
Bn2 = 2
Bn3 = 9

# simulate
tset <- seq(0, 60, 1)
l1.simu <- NaN*tset; l1.simu[1] <- l10
n1.simu <- NaN*tset; n1.simu[1] <- n10
l2.simu <- NaN*tset; l2.simu[1] <- l20
n2.simu <- NaN*tset; n2.simu[1] <- n20
l3.simu <- NaN*tset; l3.simu[1] <- l30
n3.simu <- NaN*tset; n3.simu[1] <- n30

# l10 <- l1.simu[length(tset)]
# n10 <- n1.simu[length(tset)]
# l20 <- l2.simu[length(tset)]
# n20 <- n2.simu[length(tset)]
# l30 <- l3.simu[length(tset)]
# n30 <- n3.simu[length(tset)]

B = 2
for(i in 2:length(tset)) {
  dt <- tset[i] - tset[i-1]
    
  dl1 <- N1*B*l1[i]*dt
  dn1<- N1*B*n1[i]*dt
  
  dl2<- N2*B*l2[i]*dt
  dn2<- N2*B*n2[i]*dt
  
  dl3<- N3*Bl3[i]*dt
  dn3<- N3*Bn3[i]*dt
  
  l1.simu[1] <- l1.simu[i-1] + dl1
  n1.simu[1] <- n1.simu[i-1] + dn1
  l2.simu[1] <- l2.simu[i-1] + dl2
  mn2.simu[1] <- n2.simu[i-1] + dn2
  ml3.simu[1] <- l3.simu[i-1] + dl3
  mn3.simu[1] <- n3.simu[i-1] + dn3
}


dl1 <- N1*Bl1
  dn1<- N1*Bn1
  
  dl2<- N2*Bl2
  dn2<- N2*Bn2
  
  dl3<- N3*Bl2
  dn3<- N3*Bn2

  
```

